Distribution-oblivious Programming with Dish

### 2019-09-17 pldi/draft.tex
Update preamble
Minor notes in intro

%% Title information
\title{Distribution-oblivious Programming}         %% [Short Title] is optional;
% \titlenote{with title note}             %% \titlenote is optional;
%                                         %% can be repeated if necessary;
%                                         %% contents suppressed with 'anonymous'
% \subtitle{Subtitle}                     %% \subtitle is optional
% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
\newcommand{\sx}[1]{(\S\ref{#1})}
\newcommand{\sys}{{\scshape Dish}\xspace}

@@

Distributed systems offer significant benefits over their centralized counterparts.
By parallelizing computation they can  
can speed up computations, mitigate resource-exhaustion attacks, improve fault-tolerance, and balance load during spikes.
% offer notable benefits over their centralized counterparts:
%   using partitioning and replication, they can process and store data with increased throughput and fault-tolerance.
Yet, only a minority of developers, employed by the select few companies that deal with massive datasets, have the luxury of engineering software systems with distribution baked in from the start.
The remaining majority starts by developing and deploying software in a centralized manner---that is, \emph{until} there is a significant change of requirements, such as a load increase.

Despite these benefits, the development and deployment of distributed software remains different from and significantly more difficult than its centralized counterpart.
Whereas anyone can stitch together a Bash script to compute on a single machine, 
  % domain-experts routinely glue scripts together to process and share data. % without the help of a computing expert.
   scaling out to multiple computers requires expert labor around ``point'' solutions with expensive setups, restricted programming interfaces, and exorbitant composition costs~\cite{maas2014holistic, schwarzkopf2013dios, andromeda:hotos:2015, jonas2017occupy, varghese2018next, nefele:18}.
Consider the following Unix pipeline featured in Bentley's Programming Pearls column, slightly modified to work on today's Unix.

% \begin{lstlisting}[language=sh,float=h,numbers=none]
% cat doc.ms |                   # print input file
% groff -t -e -mandoc -Tascii |  # remove formatting
% col -bx |                      # remove backspaces 
% tr A-Z a-z |                   # convert to lower case
% tr -d '[:punct:]' |            # remove punctuation
% sort |                         # put words in alphabetical order
% unique |                       # remove duplicate words
% comm -13 /usr/share/dict -     # report words not in dictionary 
% \end{lstlisting}

\begin{lstlisting}[language=sh,float=h,numbers=none]
 cat *.txt |
 tr -cs A-Za-z`\n` |
 tr A-Z a-z |
 sort |
 uniq -c |
 sort -rn |
 head 5 > output.txt
\end{lstlisting}

This pipeline takes an input file (line 1), removes

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

@@

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

>>>>

%% Title information
\title{Distribution-oblivious Programming with Dish}         %% [Short Title] is optional;
% \titlenote{with title note}             %% \titlenote is optional;
%                                         %% can be repeated if necessary;
%                                         %% contents suppressed with 'anonymous'
% \subtitle{Subtitle}                     %% \subtitle is optional
% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

\newcommand{\cf}[1]{(\emph{Cf}.\S\ref{#1})}
\newcommand{\sx}[1]{(\S\ref{#1})}
\newcommand{\sys}{{\scshape Dish}\xspace}
\newcommand{\unix}{{\scshape Unix}\xspace}

\setlist{noitemsep,leftmargin=10pt,topsep=2pt,parsep=2pt,partopsep=2pt}

@@

% 1. Shell scripting; pipelines, in particular, are a common abstraction for expressing filters
% They are an easy way to , because they combine a set of assumptions that work well with Unix
% They work great on a single machine, but are difficult to scale out
% Could we fully automate distribution? 
% The key insight is that pipelines already express a domain-specific
% computation that is easily amenable to distribution.
%

Distributed systems offer significant benefits over their centralized counterparts.
For example, they can speed up computations and can process amounts that would not fit into any single machine.
% offer notable benefits over their centralized counterparts:
%   using partitioning and replication, they can process and store data with increased throughput and fault-tolerance.
% Yet, only a minority of developers, employed by the select few companies that deal with massive datasets, have the luxury of engineering software systems with distribution baked in from the start.
% The remaining majority starts by developing and deploying software in a centralized manner---that is, \emph{until} there is a significant change of requirements, such as a load increase.

Despite these benefits, their development remains different from and significantly more difficult than their centralized counterparts.
Whereas anyone can stitch together a Bash script to compute on a single computer, 
  % domain-experts routinely glue scripts together to process and share data. % without the help of a computing expert.
   scaling out to multiple ones requires expert labor around ``point'' solutions with expensive setups, restricted programming interfaces, and exorbitant composition costs~\cite{maas2014holistic, schwarzkopf2013dios, andromeda:hotos:2015, jonas2017occupy, varghese2018next, nefele:18}.

To understand this sharp contrast, consider calculating frequencies of terms within a set of input files.
For a developer with \emph{one} computer, this is achievable with a short \unix pipeline:

% \begin{lstlisting}[language=sh,float=h,numbers=none]
% cat doc.ms |                   # print input file
% groff -t -e -mandoc -Tascii |  # remove formatting
% col -bx |                      # remove backspaces 
% tr A-Z a-z |                   # convert to lower case
% tr -d '[:punct:]' |            # remove punctuation
% sort |                         # put words in alphabetical order
% unique |                       # remove duplicate words
% comm -13 /usr/share/dict -     # report words not in dictionary 
% \end{lstlisting}
% greece; stops
% AWS policy specification
\begin{lstlisting}[language=sh,float=h,numbers=none]
 cat * | tr -cs A-Za-z`\n` | tr A-Z a-z |
   sort | uniq -c | sort -rn | head 5 > out
\end{lstlisting}

Combining several features, Unix makes small tasks easy to express;
  composition allows the composition of general primitives
Key elements of \unix are the ability to stitch together programs written in different languages, the abstraction of a file system as a set of resident streams, as well as naming---the ability to resolve command, file names within a global context.
Under the hood, the \unix kernel buffers results, concurrency control

% Key insight 
% % The garden-hose philosophy is similar to Haskell 
% This pipeline takes an input file (line 1), removes 
% * do not need to use specialized frameworks for composition---instead, as long as they conform to the shell, individual primitives can be written in any programming language.
% * do not need to rewrite ---

Unfortunately, 
A developer with many computers has to choose between three options, all of which require significant manual effort.
The most popular is to leverage domain-specific ; this requires rewriting portions of the program in a new language and use operations

% If developers have already expressed their computation as a Unix pipeline, they should not have to manually rewrite the program in other environment---\eg Hadoop, Spark---to exploit distribution.
For simple, one-off pipelines that take a few minutes to compose this cost---order of magnitude higher than---becomes unjustifiable.
For many legacy pipelines  
Could the generation and execution of a distributing version of the pipeline be automated?

The key insight behind our work is that Unix pipelines already provide most of the information for distributing a computation.
Indeed, the primary reason behind $p_1$'s succinctness is that the pipeline is a domain-specific language for describing operations on streams.
Primitives 
\sys enables \emph{distribution-oblivious programming}: 
  multi-order of magnitude speedups with correctness guarantees and without any developer effort.

The a series of transformations that 
Second, we are using describing composition . For this, we are restricting the study to a subset of the POSIX shell--

Local operations and identifiers have to be carefully translated to distributed ones, a translation that depends on the types of the operations and associated identifiers.

The semantics is enough to capture powerful features such as stream branching and feedback constructs found in practice and in the system's evaluation.

The paper is structured as follows.
We first start with a background section outlining pipeline concepts.
Sections \ref{profiling}--\ref{recipes} highlight our key contributions:
\begin{itemize}

  \item
  \S\ref{overview} overviews the system and introduces the notion of command classes.
	introduces \emph{load-detection transformations} that collect windowed statistics about load at each module boundary.
  Control-plane coordination with a global view of load and available resources helps decide when to initiate scale-out of a bottlenecked module.

  \item
  \S\ref{core} presents a set of parametrizable \emph{distribution transformations} that transparently scale modules out.
  These transformations can create module replicas, hook communication channels among them, schedule requests, and forward side-effects such as mutation and collection of memory.

  \item
  \S\ref{recipes} outlines \emph{distribution recipes}, lightweight annotations that guide the semantics of the resulting distribution.
  They offer significant flexibility by\eat{ generating configurations that} parametrizing transformations, including tuning state management, replication consistency, event propagation, and colocation preferences.
\end{itemize}

\noindent
\sys's evaluation~\sx{eval} uses a combination of one-liner micro-benchmarks that highlight certain features and three multi-line programs---a web crawling and indexing engine, a production genomics pipeline, and a climate data processing.
We close with a discussion of related prior work~\sx{related} and possible future directions~\sx{discussion}.

\section{Background}

\begin{acks}
  % Dumping people so that we don't forget
  % 
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

@@

%% Appendix
\appendix
\section{Scripts used in the evaluation}

This appendix contains the source code of the scripts used in the evaluation of
the \sys. They are part of the codebase (released as open source with the camera
ready), and are provided here only to aid the reviewers.
